---
title: "Final_Project_2"
author: "Meher Mankikar"
date: "July 12, 2019"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## R Markdown
Introduction:

In this project, I used the Iris Species data set. This data set includes data about the Iris flower's Species, Petal Length, Petal Width, Sepal Length and Sepal Width. Using this data, I aimed to use data science techniques in order to analyze the data of different Iris species. I then used this model to predict the species of flowers based on the other factors for a test set. 

Method/Analysis: 

First the Iris Species data set was loaded. 
```{r}

if(!require(readr)) install.packages("readr", repos = "http://cran.us.r-project.org")
library(readr)

iris_data <- read_csv("/Users/mehermankikar/Downloads/iris-species/Iris.csv")
iris_data
colnames(iris_data) <- c("ID", "Sepal.Length", "Sepal.Width", "Petal.Length", "Petal.Width", "Species")
str(iris_data)

```

Then, the data was split into training and tests sets. 
```{r}

library(caret)
##Making test and train sets
index <- createDataPartition(iris_data$Species, p = 0.50, list = FALSE)
iris_train <- iris_data[index,]
str(iris_train)
iris_test <- iris_data[-index,]
str(iris_test)

```

Data Visualization: 

Next, the data was examined through the summary method. A boxplot was also created to better visualize the data. 
```{r}

##Looking at the Data
summary(iris_train)

x <- iris_train[,1:4]
y <- iris_train[,5]
par(mfrow=c(1,4))
for(i in 1:4) {
  boxplot(x[i], main=names(iris_train)[i])
}

```


After analyzing the data, three models were created that would be used for prediction of the flower species later. 
```{r}

##Creating Models

library(caret)
control <- trainControl(method='cv', number=10)
metric <- 'Accuracy'
#LDA
set.seed(101)
fit.lda <- train(Species~., data=iris_train, method='rf', 
                 trControl=control, metric = metric)
#KNN
set.seed(101)
fit.knn <- train(Species~., data=iris_train, method='knn', 
                 trControl=control, metric=metric)
#RF
set.seed(101)
fit.rf <- train(Species~., data=iris_train, method='ranger', 
                trControl=control, metric=metric)

iris.results <- resamples(list(lda=fit.lda, knn=fit.knn,rf=fit.rf))

```

Results: 

Finally, the results of the three models were seen and the models were used to make predictions of flower species on the test set. The Accuracy of the model was 98.67%.
```{r}

# Results
summary(iris.results)

#Making Predictions and Testing Accuracy
iris_prediction <- predict(fit.lda, iris_test)
confusionMatrix(table(iris_prediction, iris_test$Species))

```

Conclusion: 
In this project, I successfully separated the Iris Species data set and analyzed the data. With this information, a model was created to use the different features of the flower to predict the species. The model was tested on the test set and had an accuracy of 98.67%. In the future, with a larger data set and more factors for each flower, a more accurate model could be created. 
